---
title: "Report_Exercise_8"
author: "Willy Müller"
date: "2023-06-05"
output: html_document
---





```{r}
library(rsample)
library(dplyr)
library(tidyr)
library(caret)
library(readr)


davos <- read_csv("C:/Users/Admin/Documents/agds_report_willymuller/Data/FLX_CH-Dav_FLUXNET2015_FULLSET_DD_1997-2014_1-3.csv") 
laegern <- read.csv("C:/Users/Admin/Documents/agds_report_willymuller/Data/FLX_CH-Lae_FLUXNET2015_FULLSET_DD_2004-2014_1-4.csv")

```
```{r}
set.seed(1982)
davos_split <- initial_split(davos, prop = 0.8)
davos_train <- training(davos_split)
davos_test <- testing(davos_split)

set.seed(1982)  
laegern_split <- initial_split(laegern, prop = 0.8)
laegern_train <- training(laegern_split)
laegern_test <- testing(laegern_split)
```


```{r}
mod_davos <- caret::train(
  GPP_NT_VUT_REF ~ SW_IN_F + VPD_F + TA_F,
  data = davos_train,
  method = "knn",
  trControl = caret::trainControl(method = "none"),
  tuneGrid = data.frame(k = 1),
  metric = "MAE"
)

mod_laegern <- caret::train(
  GPP_NT_VUT_REF ~ SW_IN_F + VPD_F + TA_F,
  data = laegern_train,
  method = "knn",
  trControl = caret::trainControl(method = "none"),
  tuneGrid = data.frame(k = 1),
  metric = "MAE"
)

davos_preds_within <- predict(mod_davos, newdata = davos_test)
laegern_preds_within <- predict(mod_laegern, newdata = laegern_test)

davos_preds_across <- predict(mod_davos, newdata = laegern_test)
laegern_preds_across <- predict(mod_laegern, newdata = davos_test)

metrics_within <- function(pred, actual) {
  mae <- mean(abs(pred - actual))
  rmse <- sqrt(mean((pred - actual)^2))
  r2 <- caret::R2(pred, actual)
  return(list(mae = mae, rmse = rmse, r2 = r2))
}

davos_metrics_within <- metrics_within(davos_preds_within, davos_test$GPP_NT_VUT_REF)
laegern_metrics_within <- metrics_within(laegern_preds_within, laegern_test$GPP_NT_VUT_REF)

metrics_across <- function(pred, actual) {
  mae <- mean(abs(pred - actual))
  rmse <- sqrt(mean((pred - actual)^2))
  r2 <- caret::R2(pred, actual)
  return(list(mae = mae, rmse = rmse, r2 = r2))
}

davos_metrics_across <- metrics_across(davos_preds_across, laegern_test$GPP_NT_VUT_REF)
laegern_metrics_across <- metrics_across(laegern_preds_across, davos_test$GPP_NT_VUT_REF)

cat("Within-site predictions - Davos:\n")
cat("MAE:", davos_metrics_within$mae, "\n")
cat("RMSE:", davos_metrics_within$rmse, "\n")
cat("R-squared:", davos_metrics_within$r2, "\n")

cat("\nWithin-site predictions - Laegern:\n")
cat("MAE:", laegern_metrics_within$mae, "\n")
cat("RMSE:", laegern_metrics_within$rmse, "\n")
cat("R-squared:", laegern_metrics_within$r2, "\n")

cat("\nAcross-site predictions - Davos:\n")
cat("MAE:", davos_metrics_across$mae, "\n")
cat("RMSE:", davos_metrics_across$rmse, "\n")
cat("R-squared:", davos_metrics_across$r2, "\n")

cat("\nAcross-site predictions - Laegern:\n")
cat("MAE:", laegern_metrics_across$mae, "\n")
cat("RMSE:", laegern_metrics_across$rmse, "\n")
cat("R-squared:", laegern_metrics_across$r2, "\n")

```

#Within-Site Predictions:
#The MAE and RMSE values for Davos (MAE: 1.432683, RMSE: 1.908078) are lower than those for Laegern (MAE: 2.578398, RMSE: 3.446785), indicating that the model performs better in predicting GPP  for the Davos site.
#The R-squared values for both sites indicate moderate to good model fit, with Davos having a higher R-squared value (0.5738439) compared to Laegern (0.450199).


#Across-Site Predictions:
#The MAE and RMSE values for both Davos (MAE: 2.812392, RMSE: 3.608857) and Laegern (MAE: 2.681038, RMSE: 3.418387) are higher than the within-site predictions, suggesting that the models struggle to generalize well when predicting GPP across different sites.
#The R-squared values for both sites indicate a weaker model fit compared to within-site predictions, with Davos having a slightly higher R-squared value (0.4022021) compared to Laegern (0.2871969).

#In summary, while training separate models for each site can capture site-specific relationships and result in better within-site predictions, the lower performance of the models for across-site predictions suggests limitations in generalization.




















